---
title: ""
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<h1><center><b>ROOSMAN Sales Analysis<b><center></h1>
<br>
<h2><center>------Data Cleaning--------<center></h2>

<h2>Reading Data Sets<h2>
```{r}
train_data=read.csv("Data Set Csv Rossman/train.csv")
store=read.csv("Data Set Csv Rossman/store.csv")

```

<h2> Data Set Features Description</h2>
```{r}

names(train_data)
names(store)

```

<h2> Arranging The Data Set</h2>
```{r}
library(dplyr)
train=arrange(train_data,DayOfWeek,Date)
head(train)

```
<h2>STORES  WHICH  WERE NOT OPEN HAVE NO ROLE In Analysis</h2>
```{r}
train=filter(train,Open==1)
head(train)

```

<h2>Stores Which were Open and Sales were Zero have no role in Time Series<h2>
```{r}
train=filter(train,Sales>0)
head(train)


```




```{r}
head(filter(store,is.na(CompetitionOpenSinceMonth)))

```
<h2> Merge The Train and Store Data Set in To Single Frame</h2>
```{r}
df=merge(train,store)
head(df)

```
<h2>Replacing The Null values with Median of The Column</h2>
```{r}
#For Competition Distance
m=median(df$CompetitionDistance,na.rm = TRUE)
df=df%>%mutate(CompetitionDistance=ifelse(is.na(CompetitionDistance),m,CompetitionDistance))
filter(df,is.na(CompetitionDistance))
#For Promo Week
m=median(df$Promo2SinceWeek,na.rm = TRUE)
df=df%>%mutate(Promo2SinceWeek=ifelse(is.na(Promo2SinceWeek),m,Promo2SinceWeek))
filter(df,is.na(Promo2SinceWeek))

# For Promo Year
m=median(df$Promo2SinceYear,na.rm = TRUE)
df=df%>%mutate(Promo2SinceYear=ifelse(is.na(Promo2SinceYear),m,Promo2SinceYear))
filter(df,is.na(Promo2SinceYear))
# For Competiiton Open Since Month
m=median(df$CompetitionOpenSinceMonth,na.rm = TRUE)
df=df%>%mutate(CompetitionOpenSinceMonth=ifelse(is.na(CompetitionOpenSinceMonth),m,CompetitionOpenSinceMonth))
filter(df,is.na(CompetitionOpenSinceMonth))
#For COmpetiton Open Since Year
m=median(df$CompetitionOpenSinceYear,na.rm = TRUE)
df=df%>%mutate(CompetitionOpenSinceYear=ifelse(is.na(CompetitionOpenSinceYear),m,CompetitionOpenSinceYear))
filter(df,is.na(CompetitionOpenSinceYear))

```
<h2> No Null value exist</h2>
```{r}
head(df)
```

<h2><center>------Data Visulization--------<center></h2>
<h2>CORRELATION MATRIX To See Features Importance<h2>





<h2> Sales VS Customers Relation</h2>
```{r}
library(ggplot2)
ggplot(data=df,aes(x=Customers,y=Sales))+geom_point(color="blue")

```


<h2>Sales With Respect To Promo </h2>
```{r}
sales_by_promo=df%>%group_by(Promo)%>%summarise(Average_Sales=mean(Sales))
sales_by_promo
ggplot(data=sales_by_promo,aes(x=as.factor(Promo),y=Average_Sales,fill=as.factor(Promo)))+geom_bar(stat='identity',width=0.2)+xlab("Promo")+ggtitle("Sales By Promo")


```

<h2>Sales By Prom Bases On The Week Days</h2>
```{r}

sales_by_promo=df%>%group_by(DayOfWeek,Promo)%>%summarise(Average_Sales=mean(Sales))
sales_by_promo
ggplot(data=sales_by_promo,aes(x=DayOfWeek,y=Average_Sales,fill=as.factor(Promo)))+geom_bar(stat='identity',position="dodge")

```

<h2>Sales By Month<h2>
```{r}
df=df%>%mutate(Month=format(as.Date(Date),"%m"))
sort(unique(df$Month))
sales_by_month=df%>%group_by(Month)%>%summarise(Average_Sales=mean(Sales))
sales_by_month
ggplot(data=sales_by_month,aes(x=Month,y=Average_Sales,fill=(Month)))+geom_bar(stat='identity')+ggtitle("Sales With Respect TO Month")
```

<h2> Sales With Respect To Total Days Of Month<h2>
```{r}


df=df%>%mutate(MonthDays=format(as.Date(Date),"%d"))

sort(unique(df$MonthDays))
sales_by_monthday=df%>%group_by(MonthDays)%>%summarise(Average_Sales=mean(Sales))
sales_by_monthday
ggplot(data=sales_by_monthday,aes(x=MonthDays,y=Average_Sales,fill=(MonthDays)))+geom_bar(stat='identity')+ggtitle("Sales With Respect TO Month Days")




```


<h2>Sales of Every Day in ALL MOnths</h2>
```{r}
names(df)
sales_by_month_date_days=df%>%group_by(MonthDays,Month)%>%summarise(Average_Sales=mean(Sales))
sales_by_month_date_days
ggplot(data=sales_by_month_date_days,aes(Month,Average_Sales,fill=MonthDays))+geom_bar(stat='identity')+facet_wrap(~MonthDays)

```
<h2>Sales By Week Days</h2>
```{r}

#install.packages("lubridate")
library(lubridate)
df=df%>%mutate(Week_Days=wday(as.Date(Date),week_start=1,label=TRUE))
sales_by_week_day=df%>%group_by(Week_Days)%>%summarise(Average_Sales=mean(Sales))
sales_by_week_day
ggplot(sales_by_week_day,aes(Week_Days,Average_Sales,fill=(Week_Days)))+geom_bar(stat='identity')
```
<h2>Sales BY Store Type</h2>

```{r}
sales_by_store_type=df%>%group_by(StoreType)%>%summarise(Sales=mean(Sales))
sales_by_store_type
ggplot(data=sales_by_store_type,aes(x=StoreType,y=Sales,fill=as.factor(StoreType)))+geom_bar(stat='identity')+ggtitle("Sales By Store Type")
```
<h2> Sales By Store Type and Assortment</h2>

```{r}

sales_by_assor_stype=df%>%group_by(Assortment,StoreType)%>%summarise(Average_Sales=mean(Sales))
ggplot(data=sales_by_assor_stype,aes(StoreType,Average_Sales,fill=as.factor(Assortment)))+geom_bar(stat='identity',position='dodge')

```
<h2> Find Out Which Stores Has Maximum Sales</h2>

```{r}


sales_by_store_id=df%>%group_by(Store,StoreType)%>%summarise(Average_Sales=mean(Sales))
sales_by_store_id=arrange(sales_by_store_id,desc(Average_Sales))
head(sales_by_store_id)

```
<h2>Minimum Sales</h2>
```{r}

tail(sales_by_store_id)

```
<h2>Visualization Of All Stores</h2>
```{r}
sales_by_id=df%>%group_by(Store)%>%summarise(Average_Sales=mean(Sales))
sales_by_id=arrange(sales_by_id,desc(Average_Sales))
max_stores=sales_by_id$Store[1]
min_stores=tail(sales_by_id$Store[1])

ggplot(data=sales_by_id,aes(x=Store,y=Average_Sales))+geom_point(color="blue")+
ggtitle("Average Sales BY Store Id")+
geom_text(data=sales_by_id,aes(x=max_stores,y=max(Average_Sales),label="max"))+
geom_text(data=sales_by_id,aes(x=min_stores,y=min(Average_Sales),label="min"))
                                                                                                                               
```
<h2>Sales By Year</h2>
```{r}

df=df%>%mutate(Year=format(as.Date(Date),"%y"))
unique(df$Year)
sales_by_year=df%>%group_by(Year)%>%summarise(Average_Sales=mean(Sales))
sales_by_year
ggplot(data=sales_by_year,aes(x=Year,y=Average_Sales,fill=Year))+geom_bar(stat='identity')+ggtitle("Sales With Respect To Year")

```
<h2>Sales With Respect To School Holidays</h2>
```{r}

sales_by_School_Holiday=df%>%group_by(SchoolHoliday)%>%summarise(Average_Sales=mean(Sales))
sales_by_School_Holiday
ggplot(data=sales_by_School_Holiday,aes(x=SchoolHoliday,y=Average_Sales,fill=SchoolHoliday))+
geom_bar(stat='identity')+ggtitle('The Effect of School Holidy on Sale')


```
<h2>The Effect of Schools Holiday On Every Day of Week</h2>

```{r}

sales_by_school=df%>%group_by(Week_Days,SchoolHoliday)%>%summarise(Average_Sales=mean(Sales))
head(sales_by_school)
ggplot(data=sales_by_school,aes(Week_Days,Average_Sales,fill=as.factor(SchoolHoliday)))+
geom_bar(stat='identity',position="dodge")+
ggtitle("The Effect Of Schools Holidays ON  ALL Week Days")


```
<h2>Sales With Respect To Competition Distance</h2>

```{r}


average_sales_by_Competition_Distance=df%>%group_by(CompetitionDistance)%>%summarise(Average_Sales=mean(Sales))
head(average_sales_by_Competition_Distance)
ggplot(data=average_sales_by_Competition_Distance,aes(x=CompetitionDistance,y=Average_Sales))+
geom_point(color="brown3")
```
